---
title: "SQL_Indexing"
author: "Sravani Saripalli"
date: '2022-07-13'
output: html_document
---

```{r}
library(RSQLite)

fpath = "/Users/sravanisaripalli/Downloads/databases/"
db_file = "sakila.sqlite"
# if database file already exists, we connect to it, otherwise
# we create a new database
db <- dbConnect(RSQLite::SQLite(), paste0(fpath,db_file))
```

```{r}
mydb = dbConnect(RMySQL::MySQL(),
                            dbname='sakila',
                            host='localhost',
                            port=3306,
                            user='admin',
                            password='admin')
```

```{sql connection=db}
select name from sqlite_schema where type='table' order by name;
```
```{sql connection=db}
PRAGMA index_list('CUSTOMER');
```
Query on a single table with a WHERE clause that is not the primary key but has an index; same query but no index on that column

#Quering without adding index

SQLite
```{r}
dbGetQuery(db, "EXPLAIN QUERY PLAN select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
start_time <- Sys.time()
dbGetQuery(db, "select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```
MySQL
```{r}
dbGetQuery(mydb, "EXPLAIN select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
start_time <- Sys.time()
dbGetQuery(mydb, "select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```

```{sql connection=db}
CREATE INDEX cust_ln ON CUSTOMER(last_name);
```

```{r}
dbSendStatement(mydb,"CREATE INDEX cust_ln ON CUSTOMER(last_name)")
dbSendStatement(db,"CREATE INDEX cust_ln ON CUSTOMER(last_name)")
```


Running the queries with index.
```{r}
dbGetQuery(mydb,"EXPLAIN select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
start_time <- Sys.time()
dbGetQuery(mydb,"select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```

```{r}
dbGetQuery(db,"EXPLAIN QUERY PLAN select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
start_time <- Sys.time()
dbGetQuery(db,"select first_name,last_name FROM CUSTOMER WHERE last_name='WILLIAMS'")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```

```{sql connection=db}
DROP INDEX cust_ln;
```

2. query with a join on at least three tables (no secondary indexes; no where clause other than join clause)

The below query is for finding the customer's address and city

```{r}
dbGetQuery(mydb,"EXPLAIN select C.first_name,C.last_name,A.ADDRESS,CI.CITY from CUSTOMER C INNER JOIN ADDRESS A ON C.address_id=A.address_id 
           inner join CITY CI ON A.city_id=CI.city_id LIMIT 10")
start_time <- Sys.time()
dbGetQuery(mydb,"select C.first_name,C.last_name,A.ADDRESS,CI.CITY from CUSTOMER C INNER JOIN ADDRESS A ON C.address_id=A.address_id 
inner join CITY CI ON A.city_id=CI.city_id LIMIT 10")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```


Similarly for sqlite
```{r}
dbGetQuery(db,"EXPLAIN QUERY PLAN select C.first_name,C.last_name,A.ADDRESS,CI.CITY from CUSTOMER C INNER JOIN ADDRESS A ON C.address_id=A.address_id inner join CITY CI ON A.city_id=CI.city_id LIMIT 10")
start_time <- Sys.time()
dbGetQuery(db,"select C.first_name,C.last_name,A.ADDRESS,CI.CITY from CUSTOMER C INNER JOIN ADDRESS A ON C.address_id=A.address_id inner join CITY CI ON A.city_id=CI.city_id LIMIT 10")
end_time <- Sys.time()
print(paste("time required:",(end_time-start_time)))
```

3. query on a single table with group by having a where clause that involves only the primary key column

Here I'm counting the number of actors with same last name by their primary key whith where condition that the actor_id is between 100 and 200.

```{r}
dbGetQuery(mydb, "explain select last_name, count(actor_id) actor_count 
from actor where actor_id <200 and actor_id>100 
group by last_name
order by actor_count desc, last_name")
start_time <- Sys.time()
dbGetQuery(mydb, "select last_name, count(actor_id) actor_count 
from actor where actor_id <200 and actor_id>100 
group by last_name
order by actor_count desc, last_name")
end_time <- Sys.time()
print(paste0("time required:",(end_time-start_time)))
```


```{r}
dbGetQuery(db, "select last_name, count(actor_id) actor_count 
from actor where actor_id <200 and actor_id>100 
group by last_name
order by actor_count desc, last_name")
start_time <- Sys.time()
dbGetQuery(db, "select last_name, count(actor_id) actor_count 
from actor where actor_id <200 and actor_id>100 
group by last_name
order by actor_count desc, last_name")
end_time <- Sys.time()
print(paste0("time required:",(end_time-start_time)))
```

4.one additional experiment of your choice: any query, any index use -- be creative and investigate

trying like operator
```{sql connection=db}
select DISTINCT C.first_name,C.last_name from CUSTOMER C inner join RENTAL R ON C.customer_id=R.customer_id 
INNER JOIN PAYMENT P ON P.rental_id=R.rental_id AND C.first_name LIKE 'MAR%';
```

Here I'm counting the number times Mary has taken rentals and has paid for it.

```{r}
dbGetQuery(mydb, "explain select count(C.customer_id) no_of_rentals from CUSTOMER C inner join RENTAL R ON C.customer_id=R.customer_id 
INNER JOIN PAYMENT P ON P.rental_id=R.rental_id AND C.first_name='MARY'")
start_time <- Sys.time()
dbGetQuery(mydb, "select count(C.customer_id) no_of_rentals from CUSTOMER C inner join RENTAL R ON C.customer_id=R.customer_id 
INNER JOIN PAYMENT P ON P.rental_id=R.rental_id AND C.first_name='MARY'")
end_time <- Sys.time()
print(paste0("time required:",(end_time-start_time)))
```


```{r}
dbGetQuery(db, "select count(C.customer_id) no_of_rentals from CUSTOMER C inner join RENTAL R ON C.customer_id=R.customer_id 
INNER JOIN PAYMENT P ON P.rental_id=R.rental_id AND C.first_name='MARY'")
start_time <- Sys.time()
dbGetQuery(db, "select count(C.customer_id) no_of_rentals from CUSTOMER C inner join RENTAL R ON C.customer_id=R.customer_id 
INNER JOIN PAYMENT P ON P.rental_id=R.rental_id AND C.first_name='MARY'")
end_time <- Sys.time()
print(paste0("time required:",(end_time-start_time)))
```





